Exploring probability distributions for bivariate temporal granularities

### 2019-12-17 main.Rmd
section heading changed to mapping
changing structure of smart meter application by bringing all households first

The levels of the two cyclic granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high. The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us. Plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.

@@

In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built. In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.

@@

### Interchangeability

When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets
$A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group. When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$. Although the variables used in the layers are the same, the difference in mapping leads to emphasis of different behavior of the variables.

@@

Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors. It would be interesting to explore the energy consumption distribution for these customers and gain more insights on their energy behavior which are otherwise lost either due to aggregation or looking only at coarser temporal units. The idea here is to show how looking at the time across different granularities together can help identify different behavioral patterns and identify the extreme and regular households amongst these 50 households.

@@

The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr. Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs for improving the analysis of smart meter application. The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/). This article was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all required materials required to reproduce this article.

>>>>

We will discuss the effect of the levels of cyclic granularities in this section. The levels of the two cyclic granularities plotted have an impact on the choice of plots since space and resolution might become a problem if the number of levels are too high. The criteria for different levels could be based on usual cognitive power while comparing across facets and display size available to us. Plot choices will also vary depending on which granularity is placed on the x-axis and which one across facets.

@@

We will discuss the effect of the interaction of cyclic granularities in this section. In Section \ref{sec:synergy}, we discussed how pairs of granularities can have empty combinations either due to structure of calendar, event location or duration or due to the way they are built. In this section, we will see how these empty combinations affect the visualization when a dependent variable is plotted against these granularities.

@@

### Mapping

We will discuss the effect of the mapping of cyclic granularities in this section. When we consider data sets of the form <$C_1$, $C_2$, $v$> with $C_1$ mapped to $x$ position and $C_2$ to facets $A_i$'s are placed in close proximity and each $B_j$ represent a group/facet. Gestalt theory suggests that when items are placed in close proximity, people assume that they are in the same group because they are close to one another and apart from other groups. Hence, in this case $A_i$'s are compared against each other within each group. When $C_2$ is mapped to facets and $C_1$ is mapped to $x$ axis, comparisons would ideally happen across all $B_j$ within each $A_i$. Although the variables used in the layers are the same, the difference in mapping leads to emphasis of different behavior of the variables.

@@

<!-- # how data looks -->
Smart meters provide large quantities of measurements on energy usage for households across Australia. One of the customer trial [@smart-meter] conducted as part of the Smart Grid Smart City (SGSC) project (2010-2014) in Newcastle, New South Wales and some parts of Sydney provides customer wise data on energy consumption in `general_supply_kwh` in every half hour from February 2012 to March 2014. \ref{fig:smart-meter} shows the data structure of the available data. It would be interesting to explore the energy consumption distribution for these customers and gain insights on their energy behavior which are lost either due to aggregation or looking only at coarser temporal units. The idea here is to show how looking at the energy consumption across different cyclic granularities together can help identify different behavioral patterns and identify the extreme and regular households. We take a random sample of 50 households to see how we can achieve that.

```{r smart-meter, fig.cap = "Smart meter data structure"}

smart_meter <- 
  read_rds("data/sm_cust50.rds")
```

<!-- how gravitas is used to deal with systematic exploration -->

<!-- on half-hourly energy usage and detailed information on appliance use, climate, retail and distributor product offers, and other related factors.  -->

@@

The authors would like to thank the cohort at [NUMBATS](https://www.monash.edu/news/articles/team-profile-monash-business-analytics-team), Monash University for sharing their wisdom and experience of developing R packages, Dr. Peter Toscas [from Data61 CSIRO](https://data61.csiro.au/) for providing useful inputs on improving the analysis of smart meter application. The package `gravitas` was built under [Google Summer of Code, 2019](https://summerofcode.withgoogle.com/archive/). This article was created with knitr [@knitr] and R Markdown [@rmarkdown]. The project's Github repository <https://github.com/Sayani07/paper-gravitas> contains all materials required to reproduce this article.
